---
title: "README"
author: "David Nadler, Ph.D., ASP, ENV SP"
date: "9/24/2021"
output:
  html_document: default
  pdf_document: default
email: dnadler@nyit.edu
---
```{r, echo=FALSE}
knitr::opts_chunk$set(error = TRUE)
```

# rEHS package
``` {r}

library(tidyr)
library(readr)
``` 
The **rEHS** package includes a set of tools for practical Environmental, Health and Safety management.
Download the package from: 
```{r eval = FALSE}
library(devtools)
# and install_github("nadlerphd/rEHS")
```
and open it in R:
``` {r}
library(rEHS)
```

## Calculate new concentration after a time interval 
### concentration(contaminant.generation.rate, flow.rate.cfm, timeframe, volume.cf)

**contaminant.generation.rate:** 
the rate of contaminant released in cubic feet per minute (cfm)

**flow.rate.cfm:**
the rate of air flow in cubic feet per minute (cfm)

**timeframe:**
the number of minutes (min) it takes to reach the new concentration

**volume.cf:**
the volume of the room in cubic feet (cf)

``` {r}
concentration(1.4, 2500, 11, 75000)
```
**where:**
contaminant generation is 1.4 cfm, air flow is 2500 cfm, 11 minutes is the time interval, and room is 75,000 cubic feet

***
## Calculate DART rate
### dart(dart.incidents, total.hours.worked)  

$\text{dart rate} = \frac{\text{dart incidents}}{\text{total hours worked}}$

**dart.incidents**:*
number of dart-related incidents as defined above

**total.labor.hours**
total number of hours employees worked in a reporting period

``` {r}
dart(2, 28000)
```
**where:**
for every 100 employees, 14.29 incidents resulted in lost or restricted days or job transfer due to work related injuries or illnesses

***
## Sound level at various distances
### db(dB0, distance.original, distance.new)  

$dB_{1}=dB_0+20\text{log}_{10}\frac{\text{original distance}}{\text{new distance}}$  

**dB0:**
the original sound measurement (dB)

**distance.original:**
the original distance where noise measurement was taken

**distance.new:**
the second distance to calculate the sound level reading

``` {r}
dB.distance(92, 4, 10)
```
**where:**
sound level is 92 dB at 4 feet from a source. At 10 feet from source, sound level is measured as 84.04 dB

***
## Calculate OSHA incident rate
### incident.rate(recordable.cases, total.hours.worked)  

$\text{incident rate} = \frac{\text{recordable cases}}{\text{total hours worked}}$

**recordable.cases:**
all work related deaths, illnesses, and injuries which result in a loss of consciousness, restriction of work or motion, permanent transfer to another job within the company, or that require some type of medical treatment or first-aid.

**total.hours.worked:**
total number of hours employees worked in a reporting period

``` {r}
incident.rate(17, 1500000)
```
**where:**
there are 17 recordable cases from 1.5 million hours, and for every 100 employees, 2.27 have been involved in a recordable injury or illness

***
## Conversion of mg/m^3^ to ppm
### mgm3.to.ppm(mg.per.cubic.meter, molecular.weight)  
$\text{ppm}=\frac{24.45 \times \frac{mg}{m^3}}{\text{molecular weight}}$

**mg.per.cubic.meter: ** milligrams per cubic meter

**molecular.weight:**	molecular weight

``` {r}
mgm3.to.ppm(10 ,78)
```
**where:**
concentration of acyrlamide is 10 mg/m^3^ and molecular weight of acrylamide is 78 grams/mol

***
## Conversion of ppm to mg/m^3^
### ppm.to.mgm3(ppm, molecular.weight)  
$\frac{\text{mg}}{m^3}=\frac{\text{ppm }\times \text{molecular weight}}{24.45}$  

**ppm:** concentration of molecule in parts per million

``` {r}
ppm.to.mgm3(3.13 ,78)
```
**where:** 
concentration of acyrlamide is 3.13 mg/m^3^ and molecular weight of acrylamide is 78 grams/mol

***
## Calculating the minimum air sampling volume
### min.air.volume(limit.of.quantification.mg, contaminant.target.concentration.mgm3)  

$\text{Volume}_{min}=\frac{\text{limit of quantification (mg)}}{\text{target concentration} (\text{mg/m}^{3})}$


**limit.of.quantification.mg:** the smallest amount or the lowest concentration of a substance that is possible to be determined by means of a given analytical procedure with the established accuracy, precision, and uncertainty

**contaminant.target.concentration.mgm3:** an estimate of the airborne concentrations of the contaminant being tested in mg/m^3^

``` {r}
min.air.volume(0.0027,500)
```

**where:**
a 0.0027 mg/m3 limit of quantification (concentration) exists for acetone and we are 
interested in a target concentration of 500 mg/m^3^


## Comparing an incident rate to an industry incident rate
### my.ir.comparison(my.ir, my.naics.code)

**my.ir:** input your OSHA-recordable incident rate

**my.naics.code:** input your industry's NAICS number

``` {r}
myir(2.7, 3115)
``` 

**where:**
incident rate = 2.7 and 
industry is dairy product manufacturing (NAICS 3115)

TRUE output translates to your incident rate as being lower than the industry incident rate.

***

## Calculate air flow
### q.cfm(air.velocity.fpm, area.sf)
$\text{Q}_{cfm} = \text{air velocity (fpm) }\times \text{area } (\text{ft}^{2})$

**air.velocity.fpm:** velocity of air in feet per minute

**area.sf:** cross-sectional area of air vent

``` {r}
q.cfm(5, 2)
```

**where:**
where air velocity is 5 fpm and vent cross-sectional area is 2 sf

***

## Recommended Weight Limit and Lifting Index
### rwl(horizontal.dist, vertical.dist, distance, angle, seconds.between.lifts, grasp, object.weight)


**horizontal.dist:** horizontal location of the object relative to the body (cm)

**vertical.dist:** vertical location of the object relative to the floor (cm)

**distance:** distance the object is moved vertically (cm)

**angle:**the asymmetric angle is the amount (in degrees) of trunk and shoulder rotation required by the lifting task. Angle must be input as 90, 60, 45, 30, or 0

**seconds.between.lifts:** time between lifts in seconds (s). Time must be entered as 300, 60, 30, 15, 10, or 6

**grasp:**
quality of the workers grip on the object (1 = good, 2 = fair, 3 = poor)

**object.weight:**
weight of the object in kilograms (kg)

``` {r}
rwl(45, 25, 40, 0, 15, 1, 20)
``` 

a worker is 45 cm from an object horizontally, object is 25 cm above the floor and must be lifted 40 cm up.

The worker is positioned 0 degrees relative to the 20 kg object and has a good (1) grip.

variable LI (lifting index) is created (5.328).

variable weight.limit is created (3.75).

LI interpretations:

* <=1: very low risk

* 1.01 - 1.50: low risk

* 1.51 - 2.00: moderate risk

* 2.01 - 3.00: high risk

* *> *3.01: very high risk

weight.limit interpretation:

the acceptable maximum weight (kg) that is safe under the
horizontal, vertical, and distance data input in the rwl() function,
including the angle, seconds between lifts, and grasp.

***

## Calculating the severity rate
### severity.rate(total.days.lost, total.recordable.incidents)  

$\text{severity rate} = \frac{\text{total days lost}}{\text{total recordable incidents}}$  

**total.days.lost:** 
total number of work days lost due to injury and illness. number of days lost do not need to be consecutive

**total.recordable.incidents:** total number of recordable incidents that happened in a reporting period

``` {r}
severity.rate(5, 2)
```
5 work days were lost from 2 recordable incidents  
the average recordable incident resulted in 2.5 lost days

*** 

## Permissible noise exposure
### T(dB)  

$\text{permissible exposure time}=\frac{8}{2^{0.2(\text{db - 90})}}$  

**dB:** number of decibels of noise

``` {r}
T(88)
```
**where:**
the exposure is 88 decibels

***

## Calculating the time interval in minutes for new concentration of contaminant
### time.interval(volume.cf, flow.rate.cfm, contaminant.generation.rate.cfm, concentration.ppm)

**volume.cf:** volume of room or enclosure (cubic feet)

**flow.rate.cfm:** flow rate of the air (cfm)

**contaminant.generation.rate.cfm:** rate of generation of contaminant (cfm)

**concentration.ppm:** concentration of the contaminant to predict (ppm)

``` {r}
time.interval(75000, 2500, 1.4, 175)
```

**where:**
room volume is 75,000 cubic feet
air flow rate is 2500 cfm
contaminant is generated at 1.4 cfm

***

## Calculating air velocity in feet per minute
### velocity.fpm(velocity.pressure.as.h2o)
$\text{Velocity}_{fpm}=4005\sqrt{\text{velocity pressure}}$

**velocity.pressure.as.h2o:** pressure in the direction of flow necessary to caus the air at rest to flow at a given velocity

``` {r}
velocity.fpm(.20)
```

**where:**
there is a velocity pressure of 0.2 inches

***

## Wind chill
### wind.chill(temp.fahr, wind.speed.mph)  

$\text{wind chill }F{^\circ}=35.74+0.625\text{F}^{\circ}-35.75\text{ (wind speed mph)}^{0.16} +
0.4275\text{F}^{\circ}\text{(wind speed mph)}^{0.16}$

**temp.fahr:** degrees in Fahrenheit (must be below 70)

**wind.speed.mph:** measured wind speed in miles per hour (mph)

``` {r}
wind.chill(40, 15)
```


##### **To cite this package, **
##### Nadler, David (2022). rEHS: an R package for EHS management.
